### 2024-05-12 doc/manuscript.Rmd
more lightweight analysis

---
title: "Superior paper quality with R-Markdown: The OBVIOUS Study"
bibliography: refs.bib
csl: jama.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      dpi = 600, 
                      fig.width = 11, 
                      fig.height = 6)

# load standard packages for RMD word docs
library(officer)
library(flextable)
library(gtsummary)
library(dplyr)
library(purrr)

# specify a rounding specification 
library(table.glue)

conflicted::conflict_prefer('as_flextable', winner = 'flextable')
conflicted::conflict_prefer('filter', winner = 'dplyr')

rspec <- round_spec() |>
  round_using_magnitude(digits = c(2, 1, 0),
                        breaks = c(1, 10, Inf))

# save it to options:
names(rspec) <- paste('table.glue', names(rspec), sep = '.')
options(rspec)

# default is double space 
# single spacing for title page, abstract, captions
single_par <- fp_par(line_spacing=1)

withr::with_dir(
  here::here(),
  code = {
    targets::tar_load(c(labels,
                        analysis,
                        tbl_exclusion,
                        tbl_characteristics,
                        tbl_cuminc))
  }
)

```

```{r tbl_abbrvs}

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  CVD  = 'cardiovascular disease',
  DPP  = "Diabetes Prevention Program",
  MESA = "Multi-Ethnic Study of Atherosclerosis"
)

abbrvs_paste <- function(strings){

  front <- 'Abbreviations'

  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')

  paste(front, matter, sep = ': ')

}

abbrvs_write <- function(abbr){

  sorted <- abbr[sort(names(abbr))]

  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')

  as_paragraph(abbrvs_paste(strings))

}

fn_mesa_baseline <- as_paragraph("Baseline for Multi-Ethnic Study of Atherosclerosis participants was visit 1 or 2 dependent on pre-diabetes status")

ft_tidy <- function(x){
  theme_box(x) %>% 
    align(align = 'center', part = 'all') %>%
    align(j = 1, align = 'left', part = 'all')
}

```

\newpage

Table 0: Participants included.

```{r tbl_exclusion}

tbl_exclusion %>% 
  flextable() %>% 
  width(j = 'name', width = 3) %>% 
  add_footer_lines(
    values = abbrvs_write(abbr = abbrvs[c("DPP", "MESA")])
  ) %>% 
  set_header_labels(name = "Inclusion criteria") %>% 
  footnote(part = 'body', 
           i = 2, 
           j = 'name', 
           ref_symbols = '1', 
           value = fn_mesa_baseline) %>% 
  ft_tidy()

```

\newpage

Table 1: Descriptive table of Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis participants included in the current analysis

```{r}

tbl_characteristics$included %>% 
  as_flex_table() %>% 
  ft_tidy()

```

\newpage

Table 2: Descriptive table of Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis participants excluded from the current analysis

```{r}

tbl_characteristics$excluded %>% 
  as_flex_table() %>% 
  ft_tidy()

```

\newpage

Table 3. Cumulative incidence of diabetes over follow-up in the Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis

```{r}

tbl_cuminc %>% 
  ft_tidy() %>% 
  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')

```

\newpage

Table 4.

```{r}

analysis$eval %>% 
  pivot_wider(names_from = evaluation, 
              values_from = c(standard, individualized)) %>% 
  select(stat, 
         standard_internal,
         individualized_internal,
         standard_external,
         individualized_external) %>% 
  separate(stat, into = c("group", "stat"), sep = '\\.\\.', fill = 'left') %>% 
  mutate(
    stat = recode(
      stat,
      demo_parity = "Demographic parity",
      equal_oppo = "Equal opportunity",
      equal_odds = "Equal odds"
    ),
    group = recode(
      group,
      race = "Race/ethnicity",
      sex = "Sex"
    )
  ) %>% 
  as_grouped_data(group = 'group') %>% 
  janitor::remove_empty('rows') %>% 
  as_flextable(hide_grouplabel = TRUE) %>%
  set_header_labels(
    stat = "Evaluation statistic",
    standard_internal = "Standard",
    individualized_internal = "Individualized",
    standard_external = "Standard",
    individualized_external = "Individualized"
  ) %>% 
  add_header_row(values = c("Evaluation statistic", 
                            "Internal Validation", 
                            "External Validation"), 
                 colwidths = c(1, 2, 2)) %>% 
  ft_tidy() %>% 
  width(width = 1.5) %>% 
  merge_v(part = 'header', j = 1) %>% 
  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')

```