### 2025-04-04 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# DPP-individualized-risk-prediction

<!-- badges: start -->
<!-- badges: end -->

## Reproducing

We use `targets` to coordinate this project. Results can be reproduced by opening the project and:

1. install dependencies by running `renv::restore()` in your R console.

2. run the pipeline by running `targets::tar_make()`.